---
title: "Chicago Battery Crime Report"
author: "Yi Ma, Haoxuan Sun, Chenshuo Pan, Yuning Wang"
date: '2022-12-16'
output: github_document
---

**Introduction:**\
As a UW-Madison student, Chicago is our first choice for holidays and weekends, \
so safety issues are worth our research. Our project aims to help people to \
understand the crime rate in Chicago and then plan their trips more wisely by \
analyzing the crime statistics in Chicago.\
In popular views, people commonly believe that the southern part of Chicago is \
more dangerous than the northern part. However, we think this view might be biased. \
This paper investigates public security based on crime rate with different factors \
such as location, date, crime type, etc. We use arrest probability as a standard \
to measure security in Chicago. We assume arrest probability serves as a deterrent, \
and a higher arrest rate results in a lower crime count. However, we found \
no association between the rate of change for crime count and the rate of change \
for arrest probability by month for each region. \

```{r, echo = F, warning=F, message = F,results="hide"}
library(sf)
library(ggplot2)
library(dplyr)
library("tsibbledata")
library("lubridate")
library(feasts)
library(tsibble)
library(shiny)
library(tidyverse)
library(shinythemes)
library(car)
Sys.setlocale("LC_TIME", "English")
```
**Data:**\
```{r, echo = F}
df = read.csv("crime2021.csv")
df$Date = df$Date %>% as.Date(format = "%m/%d/%Y")
df1 = df%>%
  drop_na()
df2 = df%>%
  filter(Primary.Type=="BATTERY")%>%
  mutate(month = month(Date))
lm1 = glm(as.factor(Arrest)~Domestic+as.factor(Community.Area)+as.factor(month)+
            as.factor(Description),data = df2,family = binomial())
```


```{r, fig.height=8, fig.width=8, echo = F}
chicago_geom <- read_sf('geo_export_216637c9-3ad0-40ba-86e1-3c2af0e62e79.shx')
df$Month = as.integer(substr(df$Date, 1, 2))
df$Date = df$Date %>% as.Date(format = "%m/%d/%Y")

arr<-df %>% select(c(Date, Arrest,Primary.Type, Location.Description))%>%
filter(Primary.Type == "BATTERY")%>%
filter(Arrest == "true")%>%
  group_by(Date) %>%
  summarise(arrest_number = n())

time_df = df %>% select(c(Date, Primary.Type, Location.Description))%>%
  filter(Primary.Type == "BATTERY") %>%
  group_by(Date) %>%
  summarise(count = n())%>%left_join(.,arr,by ="Date")%>%mutate(prob = arrest_number/count)

time_df = as_tsibble(time_df, index = Date)
time_df = fill_gaps(time_df, .full = FALSE)
cols = scales::viridis_pal()(10)
```
```{r, echo = F}
head(df2)
```

Our raw data is from the Chicago city government \
(https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-Present/ijzp-q8t2). \
It contains the crime incidents of murder in Chicago From 2001 to the present. \
The original data contains 7666107 rows and 22 columns. Each row in the dataset \
represents a reported crime happening in Chicago. Since the original dataset \
contains too many rows, we only selected the data of 2021 for the timeliness, \
and we selected several key columns of data, such as crime type, description, \
domestic, community area, etc. The dataset has some missing values in the column \
Longitude and Latitude. Since the official conceals the detailed block number \
due to privacy issues, it is impossible to impute the missing data. However, \
we still want to keep the rows with the missing locations since they may contain \
useful Date and Type information. We focus on the battery crime rate in Chicago \
since the number of reported battery crimes is the second highest among all \
types of crimes in the dataset. And according to the sentencing guidelines of \
Illinois, this type of crime should be arrested legally (Sentencing guidelines \
for Class A misdemeanors can be found under Illinois statute 730 ILCS 5/5-4.5-55.). \
Therefore, it is reasonable to use this probability as an indicator of safety in Chicago.\


**Method/Procedure:**\
At first, we used Logistic Regression to predict the probability of battery crime \
arrest rate with explanatory variables of the community area, month, crime \
description, and domestic status to predict the arrest rate. Considering that \
our responsive variable, “Arrest”, is a binary variable (arrest or not), logistic \
regression is a desirable method. In addition, logistic regression is also \
convenient for us to calculate the odds ratio of different conditions. In order \
to help us better observe the data in different scopes. We visualize the arrest \
probability and crime count for each community area and each month using a \
shiny app. It allows us to select certain regions and months and then create the \
map with probability statistics. The shiny app will return an arrest probability \
and two plots. Tuning community area, month, crime description, and domestic \
status will calculate the arrest rate under selected conditions. For the \
time series plot, users can select from Year, Month, and Week to visualize the \
trend of arrest rate within that period. In the map plot, users can show the \
arrest rate of all community areas within a range of two months. We can see the \
arrest rate and crime count change when the month changes given other factors stay the same. \




```{r, eval = FALSE, echo = F}
ui<-fluidPage(
  theme = shinythemes::shinytheme('superhero'),
  titlePanel("Chicago Crime"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Calculating the 2021 Chicago Arrest Rate for Different Conditions"),
      selectInput("Domestictrue", 
                  label = "Choose Domestictrue",
                  choices = c("true","false"),
                  selected = "true"),
      selectInput("Community", 
                  label = "Choose Community.Area",
                  choices = seq(1,77,1),
                  selected = "1"),
      selectInput("month", 
                  label = "Choose month",
                  choices = seq(1,12,1),
                  selected = "1"),
      selectInput("Description", 
                  label = "Choose a Description",
                  # choices = c("ATTEMPT THEFT", "DELIVERY CONTAINER THEFT",
                  #             "FROM BUILDING", "FROM COIN-OPERATED MACHINE OR DEVICE",
                  #             "$500 AND UNDER","OVER $500",
                  #             "POCKET-PICKING","PURSE-SNATCHING",
                  #             "RETAIL THEFT"),
                  
                  choices = c("AGG. DOMESTIC BATTERY - HANDS, FISTS, FEET, SERIOUS INJURY", "AGG. PROTECTED EMPLOYEE - HANDS, FISTS, FEET, SERIOUS INJURY",
                              "AGGRAVATED - HANDGUN", "AGGRAVATED - HANDS, FISTS, FEET, NO / MINOR INJURY",
                              "AGGRAVATED - HANDS, FISTS, FEET, SERIOUS INJURY","AGGRAVATED - KNIFE / CUTTING INSTRUMENT",
                              "AGGRAVATED - OTHER DANGEROUS WEAPON","AGGRAVATED - OTHER FIREARM",
                              "AGGRAVATED DOMESTIC BATTERY - HANDGUN","AGGRAVATED DOMESTIC BATTERY - KNIFE / CUTTING INSTRUMENT",
                              "AGGRAVATED DOMESTIC BATTERY - OTHER DANGEROUS WEAPON","AGGRAVATED OF A CHILD",
                              "AGGRAVATED OF A SENIOR CITIZEN","AGGRAVATED OF AN UNBORN CHILD",
                              "AGGRAVATED P.O. - HANDS, FISTS, FEET, NO / MINOR INJURY","AGGRAVATED P.O. - HANDS, FISTS, FEET, SERIOUS INJURY",
                              "AGGRAVATED POLICE OFFICER - HANDGUN","AGGRAVATED POLICE OFFICER - OTHER DANGEROUS WEAPON",
                              "AGGRAVATED POLICE OFFICER - OTHER FIREARM","AGGRAVATED PROTECTED EMPLOYEE - HANDGUN",
                              "AGGRAVATED PROTECTED EMPLOYEE - KNIFE / CUTTING INSTRUMENT","AGGRAVATED PROTECTED EMPLOYEE - OTHER DANGEROUS WEAPON",
                              "DOMESTIC BATTERY SIMPLE","OF AN UNBORN CHILD",
                              "PROTECTED EMPLOYEE - HANDS, FISTS, FEET, NO / MINOR INJURY","SIMPLE"),
                  
                  
                  selected = "RETAIL THEFT"),
   sliderInput("Month",
               label = "Select a month",
               min = 1, max = 12, value = c(1, 12)),
   
         titlePanel("Time Series"),
   selectInput("Date",
               label = "Choose a variable to display",
               choices = c("year",
                           "month",
                           "week"),
               selected = "year")

      
      

    ),
    mainPanel(
      h3(textOutput("text1")),
      
      plotOutput("map_plot"),
      plotOutput("bar_plot"),
      plotOutput("ts_plot")

    )
  )
)

server <- function(input, output) {
    predict.data = reactive({data.frame(Domestic = input$Domestictrue,Community.Area = input$Community
                                ,month = input$month,Description = input$Description)})
    linear.pred.1 = reactive({predict(object = lm1,
                            newdata = predict.data(), type = "response")})
    
    output$text1 <- renderText({ 
      paste("probability is", linear.pred.1())
    })
    output$ts_plot <- renderPlot({

      gg_season(time_df, prob, period = input$Date, pal = cols) + 
        theme_bw()

    })
    output$map_plot <- renderPlot({
      
      count = df2 %>% filter(Primary.Type == "BATTERY", (as.integer(month) >= input$Month[1]) & (as.integer(month) <= input$Month[2])) %>% 
        select(Community.Area, month) %>% 
        group_by(Community.Area) %>% 
        summarise(count = n())
      
      arr<-df2 %>% filter(Primary.Type == "BATTERY", (as.integer(month) >= input$Month[1]) & (as.integer(month) <= input$Month[2])) %>% 
        filter(Arrest == "true") %>%
        select(Community.Area, month) %>% 
        group_by(Community.Area) %>% 
        summarise(arrest_number = n()) 
      
      probability = left_join(count,arr,by = 'Community.Area')%>%mutate(prob = arrest_number/count)

      chicago_geom$area_num_1 = as.integer(chicago_geom$area_num_1)
      
      chicago_geom %>% left_join(probability, by = c("area_num_1" = "Community.Area"))%>% 
        ggplot() + 
          geom_sf(size = 0.2, color = "black", aes(fill = prob)) + 
          ggtitle("Battery Crime arrest probability  Plot") + 
          coord_sf() + 
          scale_fill_viridis_c(option = "magma",begin = 0.1) + 
          theme_bw()

    })
    output$bar_plot <- renderPlot({
      
        arr<-df2 %>% filter(Primary.Type == "BATTERY", (as.integer(month) >= input$Month[1]) & (as.integer(month) <= input$Month[2])) %>% 
          filter(Arrest == "true") %>%
          select(Community.Area, month) %>% 
          group_by(Community.Area) %>% 
          summarise(arrest_number = n())
        
        count = df2 %>% filter(Primary.Type == "BATTERY", (as.integer(month) >= input$Month[1]) & (as.integer(month) <= input$Month[2])) %>% 
          select(Community.Area, month) %>% 
          group_by(Community.Area) %>% 
          summarise(count = n())
        
        chicago_geom$area_num_1 = as.integer(chicago_geom$area_num_1)
        rate_df =  arr %>% 
          left_join(., select(chicago_geom, area_num_1, community), by = c("Community.Area" = "area_num_1")) %>% 
          left_join(., count, by = "Community.Area") %>% 
          select(community, arrest_number) %>% 
          arrange(., desc(arrest_number)) 
        rate_df[1:10,] %>% ggplot() + 
          geom_bar(aes(arrest_number, reorder(community, arrest_number), fill = arrest_number), width = 0.8, stat = "identity") + 
          theme_bw() + 
          scale_x_continuous(expand = c(0,0)) + 
          labs(x = "Arrest Number", y = "Community Area", fill = "Arrest Number")
      
    })
    
  }


app <- shinyApp(ui, server)
app
```
To see if arrest rate and crime count are correlated, we also compute the \
difference between the arrest probability and the number of crimes for the \
concurrent month and the following month in order to see if the positive change \
in arrest rate can result in a decrease in the number of crimes. We then use \
linear regression to see how those changes are correlated in order to determine \
if the arrest rate can predict future crime numbers. \




**Result:**\
The LRT for logistic regression shows that all the features are significant. \
Therefore, we can conclude that at least some of the levels in our categorical \
variables can affect the probability of arrest. We are able to apply our model \
to estimate the arrest rate in Chicago. \
For example, when the community area is 69, given other factors are constant, \
the odds of being arrested are 0.664 times of the community area is 1. When the \
time is April, given other factors are constant, the odds of being arrested are \
0.025 times of month in January. \

```{r, echo = F}
#Following two plots come from code snippets copied from our ShinyApp.
count = df2 %>% filter(Primary.Type == "BATTERY", (as.integer(month) >= 1) & (as.integer(month) <= 12)) %>% 
        select(Community.Area, month) %>% 
        group_by(Community.Area) %>% 
        summarise(count = n())
      
      arr<-df2 %>% filter(Primary.Type == "BATTERY", (as.integer(month) >= 1) & (as.integer(month) <= 12)) %>% 
  filter(Arrest == "true")%>%
        select(Community.Area, month) %>% 
        group_by(Community.Area) %>% 
        summarise(arrest_number = n()) 
      
      probability = left_join(count,arr,by = 'Community.Area')%>%mutate(prob = arrest_number/count)

      chicago_geom$area_num_1 = as.integer(chicago_geom$area_num_1)
      
      chicago_geom %>% left_join(probability, by = c("area_num_1" = "Community.Area"))%>% 
        ggplot() + 
          geom_sf(size = 0.2, color = "black", aes(fill = prob)) + 
          ggtitle("Battery Crime arrest probability Plot") + 
          coord_sf() + 
          scale_fill_viridis_c(option = "magma",begin = 0.1) + 
        theme_bw()
```

```{r, echo = F}
gg_season(time_df, prob, period = 'year', pal = cols) + 
  theme_bw()
```

In our shiny app,  we found at the beginning \
and the end of the year, there was an increase in the crime number and arrest \
probability. There exists a decreasing trend for battery arrest rate from the \
Chicago northwest community area to the southeast community area. Besides these \
two points, the battery crime arrest rate is associated with domestic types; \
The average probability of domestic battery arrest rate is 10% higher than the \
probability of non-domestic battery arrest rate.\
It is worth noting that we observe large fluctuations in the number of crimes \
during different months of the year.\
So we investigated if arrest probability is associated with crime count, \
we grouped the data by month and community area.  We fitted a linear regression \
with the rate of change in crime count as a response and the rate of change in \
arrest probability as an explanatory variable. \
The summary table shows a weak significance for rate change in arrest probability \
as the p-value is between 0.05 to 0.1. The coefficient is -0.041, it suggests \
that as one unit increases rate change in arrest probability, the number of \
crimes will decrease 0.041 units.\
However, we found no strong association between them, as the scatter plot \
between the two rates of change has a fairly random pattern, and the model has \
minimal R-squared values(0.0027).  Thus, the arrest rate cannot predict future \
crime counts in the short term accurately from our analysis so far. \
\
```{r, echo = F}
time2 = df %>% 
  filter(Primary.Type == "BATTERY") %>%
   mutate(month = month(Date), Arrest = ifelse(Arrest == "false", 0,1))%>%
  group_by(month,Community.Area) %>%
  summarise(crime_count = n(),arrest_num = sum(Arrest),arrest_prob  = arrest_num/crime_count)
time2
```

```{r, echo = F}
func <- function(x, n=1) {
  c(rep(NA, n), diff(x, n) / head(x, -1*n)*100)
}
time2$RateOfChange_crime_count <- ave(time2$crime_count, factor(time2$Community.Area), FUN=func)
time2$prob_change <- ave(time2$arrest_prob, factor(time2$Community.Area), FUN=func)
time2 = time2%>%
  mutate_if(is.numeric, function(x) ifelse(is.infinite(x), 0, x))
```

```{r, echo = F}
ggplot(data = time2) + 
  geom_point(aes(crime_count, arrest_prob)) + 
  xlab("Crime Count") + 
  ylab("Arrest Probability") + 
  theme_bw()
```


```{r, echo = F}
lm2 = lm(RateOfChange_crime_count ~ prob_change, data = time2)
summary(lm2)
```
```{r, echo = F}
Anova(lm1)
```
**Conclusion:**\
From our analysis, we generated a logistic regression model that can predict the \
arrest rate based on community area, month, crime description, and domestic status. \
For example, according to our model, the crimes involving minor injuries to \
police officers had the highest arrest probability of 0.948 in the O'hare area in January.\
\
However, we fail to conclude there’s a strong correlation between the rate of \
change in crime number and the rate of change in arrest probability, which \
indicates higher arrest rate fails to result in lower crime counts.\
Weakness: Our data set only records the cases reported in Chicago in 2021, so \
there may be cases that occurred but were not recorded in the data set. \
Furthermore, police may not arrest a suspect due to insufficient criminal \
evidence. The changes in policy may also affect the crime counts. And the \
increase in arrest rate may not affect the number of crimes immediately for \
the following month. Those factors may reduce the accuracy of our analysis. \
\
**Future Analysis:**\
We can use our model to predict the arrest rate for battery crime to determine \
the security concern for each community region in further years if months and \
specific types of crime are given. At the same time, we can also study whether \
the arrest rate can be a deterrent to crime when the time span is extended.\

Github link:[https://github.com/cspan6/STAT-433-assignment]