---
title: "421-assignment1"
author: "Chenshuo Pan"
date: '2022-09-21'
output: github_document
---

```{r}
library(dplyr)
library(tidyverse)
# install.packages("nycflights13")
library(nycflights13)
library(ggplot2)
library(mice)
library(stringr)
library(tsibbledata)
library(tsibble)
library(lubridate)
library(ggpubr)
```

```{r}
flights = nycflights13::flights
```



#1. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
flights%>%
  filter(is.na(dep_time))%>%
  summarize(n = n())
```
**8255 flights have a missing dep_time**

```{r}
#md.pattern will help us count the number of missing values in each column
md.pattern(flights)
```



**For each missing dep_time line, dep_delay,arr_time, arr_delay are all missing.**

**These rows represent the flights been cancelled.Because these lines are missing all arrival/departure** **times and expected arrival/departure times at the time.**


#2.
```{r}
convertTime <- function(data,time){
  colomn <- data$time
  time_in_minute <- column%%100 + (column%/%100)*60
  return(time_in_minute)
}
```


Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flight = flights
flight%>%
  mutate(dep_time = dep_time%%100 + (dep_time%/%100)*60,
         sched_dep_time = sched_dep_time%%100 + (sched_dep_time%/%100)*60)
```
3.Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()
```{r}
# #different line
# cancelled <- flights %>% 
#   filter(is.na(dep_delay), is.na(arr_delay))
# 
# 
# cancelled_perday<-cancelled%>%group_by(year,month,day)%>%summarize(cancel_number=n())
# 
# cancelled_perday%>%ggplot(aes(x= ,y=)) + geom_point()
# 
# per_day = flights%>%group_by(year,month,day)%>%summarize(flights_perday = n())
# 
# 
# 
# combined_per_day<-right_join(cancelled_perday,per_day)%>%mutate(proportion = cancel_number/flights_perday)
# 
# delay_perday <-flights%>%filter(!is.na(arr_delay))%>%
#   mutate(on_time = if_else(condition = arr_delay<0,
#                            true = T,
#                            false = F,
#                            missing = F))%>%
#   group_by(year,month,day)%>%summarize(average_delay = mean(on_time))
# 
# full_combined<-right_join(combined_per_day,delay_perday)


```



```{r,warning=FALSE,message=FALSE}
#one line
flight%>%filter(is.na(dep_delay), is.na(arr_delay))%>%
  group_by(year,month,day)%>%summarize(cancel_number=n())%>%
  #above codes count how many flights cancelled per day
  #Below codes count how many flights per day and combine these two data , the create a new column which
  #calculate the proportion
  right_join(flights%>%group_by(year,month,day)%>%summarize(flights_perday = n()))%>%
  mutate(proportion = cancel_number/flights_perday)%>%
  right_join(flights%>%filter(!is.na(arr_delay))%>%
  #Set the conditions(same as what we do in class) to filter the number of delayed flights per day and the total number of flights to get the average delay
  mutate(on_time = if_else(condition = arr_delay>0,
                           true = T,
                           false = F,
                           missing = F))%>%
  group_by(year,month,day)%>%
    summarize(average_delay = mean(on_time)))%>%
  replace(.,is.na(.),0)%>%
  arrange(year,month,day)%>%
  #Convert to a time series suitable for the x-axis
  mutate(date = ymd(str_c(year,month,day,sep = "-")))%>%
  as_tsibble(.,index = date)%>%
  ggarrange((ggplot(.) + geom_point(aes(x=date ,y=cancel_number))+ggtitle("cancel flights per day")+xlab("date")+ylab('number')),
            (ggplot(.)+geom_point(aes(x = average_delay, y = proportion),color="red")+ggtitle("relationship between average delay and proportion of cancelld flight per day ")+xlab("average delay")+ylab("proportion")),ncol = 2,nrow = 2)
  
  # ggplot() + geom_point(aes(x=date ,y=cancel_number))+ 
  # geom_point(aes(x = date,y = proportion),color = "red")+
  # geom_point(aes(x = date,y = average_delay),color = "blue")+ggtitle("average delay,proportion,cancel number of each date ")
```

**No apparent pattern in the number of flights cancelled per day**

**Through observation, it is not difficult to find that when the average delay is low, the proportion value is also very low, and it rises as the average delay increases. Perhaps there is a curve that can illustrate the relationship between the two**







